### 2015-04-09 reply_referee2.txt
wrote response to referees

Reply to Referee 2

Thank you for the careful and constructive review of my paper.  I have made a major revision, including a complete overhaul of the methodology, and a rewriting of most of the sections based on your comments, as well as those of the other referees.  I have also substantially shortened and at some places simplified the analysis in order to comply with Energy Journal length limits, as well as to focus on the main results.  I give a short summary of the major changes below, and then follow with a point-by-point response to your comments.

Methodologically, I have tried to reduce the complexity of the modeling by excluding the smallest of the fields - chosen as under 3 mill sm3 in recoverable reserves since these show significantly different production profiles than larger fields.  This should not have a major impact on the implications of the paper since they account for only .5 % of total recoverable oil reserves. I also exclude the Ekofisk field, as this has a unique double-hump profile, related to its early history (see p.9).

I was then able to in some ways simplify and extend the methodology. Significantly, I now use a single-variable smoothed function to control for the profile. This can be interpreted directly, and thus adds intuition to the analysis.  Perhaps more importantly, i was then able to add random effects terms - modeling the idiosyncratic variance between the fields (p. 14).  I compare these results with a model that includes a two-dimensional smoothed function (p.14-15) (but where adding a random effects term is not possible) and the main results generally agree.

To deal with the potential that the peak of production could be endogenous to price, I now also include only a single smoothed function for the full production profile, rather than splitting the function into two pieces before and after the peak.  I have further extended the model to try to deal with the issue of technological change, as well as the possibility that costs are correlated with price, which I discuss below.

The results are on the whole similar, though differ in detail. I still find no significant evidence of a contemporary effect of oil prices on production, however now I define contemporary as the same year.  In lags 1-4, modest, but significant effects are found, with the highest effect found after 2 years.

I have chosen to limit my results to the total average effect of prices on production.  Partly this is out of purely practical consideration of keeping the exposition within Energy Journal limits. But in addition, the results for different phases of production are considerably less robust, mainly because I am only using about half the data for each analysis, but also because I now am using a single smoothed function to model the production profile. Estimations of the effect of prices on investment probably deserves a seperate paper and I have also cut this section. The analysis certainly does lose some detail with these omissions, but I believe the trade-off of better clearity and robustness, and appropriate length is worth it.

Response to referee 2 comments:

"1) My greatest objection is that the paper lacks a discussion of why the author chooses to analyze a regression function as in e.g. Eq. (6) in the paper. The author has to revise this specification...To include investment and/or production costs is crucial as an increase in (marginal) costs will lead to declined production for a given oil price.

I think this is good advice, and I have redone the specifications. Significantly, I have added a new variable to account for costs.  From the year 2000 to the present this is the producer price index for oil and gas extraction from Statistics Norway. This takes into account the skyrocketing costs for rigs, steel, etc, that you note.  Unfortunately, the series does not extend farther back, so between 1960 and 2000 I use an index of US oil and gas industry costs as a proxy, with the rationale that the industry is global and costs tend to be highly correlated across regions. (p.9-10)

"Here I will add that I do not understand the sentence on p.31/32:”..a trend of increased total extraction estimates from the Norwegian continental shelf as a whole as well as from existing fields over the last 15 years of strongly rising oil prices”."

I agree, this was poorly worded - I have rewritten the entire section.

"Failure to include the demand side might be misleading, as price is a function of both demand and supply..."

I think you are right that I needed to do a better job considering endogeneity of demand in the paper.  Since prices are established in equilibrium, a bias could result if demand also was correlated with changes in field production other than through the direct price-mechanism. For example, higher GDP could mean easier lending to petroleum companies, who then might invest and increase petroleum independent of higher oil prices. I investigated this by including a measure of world GDP growth from the world bank in the regression. This did not have any effect on the estimated coeffients on price terms or on the overall fit of the model, suggesting that demand-effects are unlikely causing a bias, and therefor I excluded this variable from the final model.  I have added discussion of this on pages p. 15-16.

Other major comments:

The author leaves diagnostics and robustness checks in figures and tables in Appendix A. However, I would like a discussion in words in the text.

"I have simplified and brought forward all the analysis into the main body of the text.  I hope this will make the article easier to read and follow."

"The text is too lengthy, e.g. Fig. 3 and 4 in Section 3 are not needed, Further, above all the text in Section 5, 6 and 7 is too lengthy and repetitive."

I have eliminated some of these figures and joined together several others, and cut several sections. The manuscript is now at 25 pages plus an online appendix that details the monte-carlo experiment.

The author has to work with the language as many phrases are unfinished; a small example on p. 38: “...allowing for increased recovery percentage of the available reservoir”:

Good point - I have tried to make the exposition clearer and more concise.

1) Income tax now is 27 % (and not 28 % as is written on p. 15), while resource tax is 51 % (and not 50 % as is written on p. 15).

Thanks for the tip - I have changed the numbers in the text.