### 2016-02-21 oil_prices_title.tex
pre-OBES commit

% oil_prices_title.tex

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{natbib} %for citet and citep
\usepackage{syntonly}
\usepackage{esdiff} %for writing partial derivatives
\usepackage{url} %for inserting urls
\usepackage{placeins} % for limitting floats
\usepackage{comment} %for excluding figures
% \excludecomment{figure} % for printing without figures
% \syntaxonly % for quickly checking document
%set document settings

\title{The Effect of Oil Price on Field Production: Evidence from the Norwegian Continental Shelf}
\author{Johannes Mauritzen\\
		Department of Industrial Economics and Technology Management\\
        Norwegian University of Science and Technology\\
        7491 Trondheim
        Norway\\
        johannes.mauritzen@iot.ntnu.no\\
        \url{jmaurit.github.io}\\
		}
\date{\today}

\begin{document}
	\maketitle

\begin{abstract}
I use detailed field-level data on Norwegian offshore oil production and a semi-parametric additive model to control for the production profile of fields to estimate the effect of oil prices on production.  I find no significant evidence of a concurrent reaction of field production to oil prices, though a slight lagged effect is found of the magnitude of approximately 1 to 4\% for a 10 dollar per barrel increase in the real price of oil.\\
\textbf{Keywords}: Oil production,Oil Prices, Norway, Generalized Additive Model.\\
\textbf{Words}: 5970
\end{abstract}

\end{document}